---
title: "Data 607 Project 4"
author: "Jeff Nieman"
date: "April 21, 2016"
output: html_document
---
##No SQL Migration

####1.  I began by loading the needed packages and calling the data in a .csv format from the github location where I stored it.  I also confirmed the dimensions of the file.

```{r, warning=F, message=F}
require(rmongodb)
require(jsonlite)
require(RCurl)
airportsurl <- getURL("https://raw.githubusercontent.com/jeffnieman11/Data607_Project4/master/airports.csv")
airports <- read.csv(text=airportsurl)
dim(airports)
head(airports)

```

####2.  I then created the mongo connection and tested to see if it is working properly.  A value of "TRUE" confirmed it.  
```{r}
mongonew <- mongo.create()
mongonew
mongo.is.connected(mongonew)

```
####3.  I then created names for a database and a collection where we add the records.  
```{r}
db <- "test.flights"
mnew <- "test.flights.airports"

mongo.get.database.collections(mongonew, db)

```
####4.  Following the advice on Andy Catlin's  to Mongo I used lappy to the split function and converted the csv file to JSON and then to bson which is the format needed for rmongodb.  A test of 3 rows confirms the approach.

```{r}

airports.1 <- lapply(split(airports, 1:nrow(airports)), function(x) mongo.bson.from.JSON(toJSON(x)))

airports.1[1:3]
```
####5.  Using the mongo.insert.batch command I inserted the data from step 4 into my mongoDB and confirmed that there indeed were the same number of records as in the csv file.
```{r}

mongo.insert.batch(mongonew, mnew, airports.1)

```

####6.  I concluded by testing several queries.  I first checked the count and records of my local airport DTW, finding to no surprise one record.  Then I tested the altitude of 645 and again just pulled up the airport of DTW.  I then queried the count and a few records at sea level, followed by queries for >=2000 feet and <0.

```{r}
mongo.count(mongonew, mnew, query ='{"faa":"DTW"}')
mongo.find.all(mongonew, mnew, query = '{"faa":"DTW"}')
mongo.find.all(mongonew, mnew, query = '{"alt":645}')
mongo.count(mongonew, mnew, query = '{"alt":0}')
mongo.find.all(mongonew, mnew, query = '{"alt":0}', limit=5)
mongo.count(mongonew, mnew, query = '{"alt":{"$gte":2000}}')
mongo.find.one(mongonew, mnew, query = '{"alt":{"$gte":2000}}')
mongo.count(mongonew, mnew, query = '{"alt":{"$lt":0}}')
mongo.find.all(mongonew, mnew, query = '{"alt":{"$lt":0}}')

```


####Conclusions:  Quite honestly I am more familiar with relational databases than I am with the noSQL database approach.  The pros of the relational database approach has to do with the structure - the columns and rows.  However the cons result the large space and process time complex and large relational databases take up.  The NoSQL approach solves this problem, but involves learning a new way to approach the queries.  Through this homework I learned how to put data into MongoDB and after a little research found the queries to be useful as well.